Unrecognized States: A Theory of Self-Determination and Foreign Influence

### 2015-10-21 draft.tex
Shortened uncertainty section and merged with outside interactions.

\section{Extensions}
\label{sec:ext}
\subsection{Introducing Uncertainty}
The assumption of perfect information may be somewhat unrealistic, and here we explore adjusting the model to accommodate some uncertainty on the part of $c$ and $p$. Payoffs as described in the stage game are those perceived by $g$ and $s$. If these payoff values are not precisely known by $c$ and $p$, enough uncertainty may be present in order for both those players to contribute resources in equilibrium.

Here, $c$ and $p$ observe a random draw of the stage game payoffs, with each payoff drawn independently and with a mean matching the original stage game payoffs. After viewing these (uncertain) payoffs, $c$ and $p$ each invest some level of resources, altering the payoff structure before it is observed (accurately) by $s$ and $g$. Based on this altered payoff structure, $c$ and $g$ choose their strategies. Uncertainty can lead to outcomes where either $c$ or $p$ invests too much, wastes resources, or makes a more severe misstep, such as $p$ investing too little and allowing reunification to occur.

In practice we expect that uncertainty is lower for the patron than for $c$ because the patron is close to, and intimately involved in, the conflict and therefore may have a better grasp of the two players' payoffs than do other outside states or groups. This makes over-contribution by $c$ more likely than under-contribution by the patron. %Of particular note here are the resources the European Union and its members have devoted toward promoting a negotiated reunification of the Turkish Republic of Northern Cyprus into Cyprus.

In the equilibria constructed in Section~\ref{sec:main}, full information on all sides implies that one of the parties would, in equilibrium, not give any resources. By adding some uncertainty about payoffs, we can account for the observed fact that $c$ sometimes expends resources unsuccessfully. This type of spending can also be explained as non-strategic spending -- i.e. spending aimed at goals other than promoting settlement, like pure humanitarianism.

@@

\subsection{Outside Interactions Between the Patron and \texorpdfstring{$c$}{c}}
The game that we model, with the control of the unrecognized state at stake, is only one of several strategic games in which the patron and $c$ may be interacting at any given time, and linkages between games are possible. We do not model any direct exchange of resources or imposition of harm between $c$ and the patron, but we take the implications of these possible outside interactions seriously. The willingness and ability of either party to contribute resources within the game we model may be affected by their interaction with one another in other contexts. As we will discuss in greater detail below, in a number of cases a patron has withdrawn support for an unrecognized state in response to pressure exerted by actors in the international community ($c$) in other venues.

@@

\section{Conclusions}
In this paper we establish unrecognized states as an outcome of interest for social scientists and provide a unified framework for analyzing that outcome and its alternatives. Current events, particularly, Russia's military and economic support of secessionist rebels in Eastern Ukraine, suggest that the phenomenon of unrecognized statehood will not fade from relevance soon. While the Ukrainian situation remains fluid, the status quo equilibrium we model is a plausible future. The equilibrium could emerge as follows: with Russian support, the secessionists solidify \emph{de facto} control of parts of Eastern Ukraine, but risk of direct confrontation with the West is sufficiently high to prevent Russian annexation of Eastern Ukraine.\footnote{Annexation of Crimea was achieved at a reasonable cost to Russia. The costs of annexation Eastern Ukraine would likely be substantially higher.} Further Russian economic aid keeps the status quo payoffs of the secessionists sufficiently high that they decline negotiated reunification. If necessary, contributions to the home state by U.S. and the EU (player $c$) strengthen the home state enough to prevent the secessionists from overthrowing the Ukrainian government and forcing recognition. In twenty years, Eastern Ukraine may look a good deal like South Ossetia does today.

>>>>

\section{Extensions}
\label{sec:ext}
%\subsection{Introducing Uncertainty}
%The assumption of perfect information may be somewhat unrealistic, and here we explore adjusting the model to accommodate some uncertainty on the part of $c$ and $p$. Payoffs as described in the stage game are those perceived by $g$ and $s$. If these payoff values are not precisely known by $c$ and $p$, enough uncertainty may be present in order for both those players to contribute resources in equilibrium.

%Here, $c$ and $p$ observe a random draw of the stage game payoffs, with each payoff drawn independently and with a mean matching the original stage game payoffs. After viewing these (uncertain) payoffs, $c$ and $p$ each invest some level of resources, altering the payoff structure before it is observed (accurately) by $s$ and $g$. Based on this altered payoff structure, $c$ and $g$ choose their strategies. Uncertainty can lead to outcomes where either $c$ or $p$ invests too much, wastes resources, or makes a more severe misstep, such as $p$ investing too little and allowing reunification to occur.

%In practice we expect that uncertainty is lower for the patron than for $c$ because the patron is close to, and intimately involved in, the conflict and therefore may have a better grasp of the two players' payoffs than do other outside states or groups. This makes over-contribution by $c$ more likely than under-contribution by the patron. %Of particular note here are the resources the European Union and its members have devoted toward promoting a negotiated reunification of the Turkish Republic of Northern Cyprus into Cyprus.

%In the equilibria constructed in Section~\ref{sec:main}, full information on all sides implies that one of the parties would, in equilibrium, not give any resources. By adding some uncertainty about payoffs, we can account for the observed fact that $c$ sometimes expends resources unsuccessfully. This type of spending can also be explained as non-strategic spending -- i.e. spending aimed at goals other than promoting settlement, like pure humanitarianism.

@@

\subsection{Uncertainty and Outside Interactions Between the Patron and \texorpdfstring{$c$}{c}}
While we present a model with perfect information, there are some empirical cases in which it appears that various players have acted on mistaken beliefs about the strength or payoffs of other players. Modeling uncertainty falls outside the scope of this paper, but uncertainty does offer an additional plausible explanation for off-equilibrium-path behavior that we observe in empirical cases.  For example, uncertainty regarding the payoff structure facing the secessionists may explain failed and costly efforts by the European Union and its members (i.e. player $c$) to promote a negotiated reunification of the Turkish Republic of Northern Cyprus into Cyprus.

The game we model, with the control of the unrecognized state at stake, is only one of several strategic games in which the patron and $c$ may be interacting at any given time, and linkages between games are possible. We do not model any direct exchange of resources or imposition of harm between $c$ and the patron, but we take the implications of these possible outside interactions seriously. The willingness and ability of either party to contribute resources within the game we model may be affected by their interaction with one another in other contexts. As we will discuss in greater detail below, in a number of cases a patron has withdrawn support for an unrecognized state in response to pressure exerted by actors in the international community ($c$) in other venues.

@@

\section{Conclusion}
In this paper we establish unrecognized states as an outcome of interest for social scientists and provide a unified framework for analyzing that outcome and its alternatives. Current events, particularly, Russia's military and economic support of secessionist rebels in Eastern Ukraine, suggest that the phenomenon of unrecognized statehood will not fade from relevance soon. While the Ukrainian situation remains fluid, as of October 2015, the status quo equilibrium we model is a plausible future. The equilibrium could emerge as follows: with Russian support, the secessionists solidify \emph{de facto} control of parts of Eastern Ukraine, but risk of direct confrontation with the West is sufficiently high to prevent Russian annexation of Eastern Ukraine.\footnote{Annexation of Crimea was achieved at a reasonable cost to Russia. The costs of annexation Eastern Ukraine would likely be substantially higher.} Further Russian economic aid keeps the status quo payoffs of the secessionists sufficiently high that they decline negotiated reunification. If necessary, contributions to the home state by U.S. and the EU (player $c$) strengthen the home state enough to prevent the secessionists from overthrowing the Ukrainian government and forcing recognition. In twenty years, Eastern Ukraine may look a good deal like South Ossetia does today.
